---
title: "DATA 606 Data Project Proposal"
author: "null"
output: pdf_document
---

### Data Preparation

```{r}
library(tidyverse)
library(lubridate)
library(dplyr)
library(Hmisc)
```




```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data
WorksiteMeetings <- read.csv("WorksiteMeetings.csv") 
```

```{r}
glimpse(WorksiteMeetings)
```

```{r}
summary(WorksiteMeetings)
```

```{r}
WorksiteMeetings$datecreated <- mdy(WorksiteMeetings$datecreated)
```

```{r}
WorksiteMeetings$memberCount <- as.numeric(WorksiteMeetings$memberCount)
WorksiteMeetings$ADFcount <- as.numeric(WorksiteMeetings$ADFcount)
WorksiteMeetings$turnoutcount <- as.numeric(WorksiteMeetings$turnoutcount)
```
change the NA's in ADF to zeros, i left the memberCounts as NA bc i dont know how many it really should be
```{r}
WorksiteMeetings$ADFcount <- WorksiteMeetings$ADFcount %>% replace(is.na(.), 0)

```

**New fields for turnout percentage and ADF percentage of building**
**POSSIBLE PROBLEM: there are 4 buildings with a percent higher than 100%, this could happen if the roster size changed since the event happened. these are all in smaller buildings, i think i will just count them as 100%, as they're off by 1 or 2 and they got the entire size of the current building to come to event. this could be an issue with other buildings that i cant detect under 100%, but i dont have a better way to compare each individual roster size at that time of the event, so i'm using the current roster as a proxy, building rosters ~usually~ stay around the same size**
```{r}
WorksiteMeetings$ADFPercentage <- (WorksiteMeetings$ADFcount / WorksiteMeetings$memberCount)
WorksiteMeetings$TurnoutPercentage <- (WorksiteMeetings$turnoutcount / WorksiteMeetings$memberCount)
```

Replace those >100% with 100%
```{r}
WorksiteMeetings$TurnoutPercentage <-replace(WorksiteMeetings$TurnoutPercentage, WorksiteMeetings$TurnoutPercentage>1, 1) 

```

Remove the NY Security, there are too many huge memberCount accounts of grouped locations. this is a known issue, os im unsurprised to see it here. It makes it really hard to figure out which buildings our members actually work at:
```{r}
WorksiteMeetings <- WorksiteMeetings %>%  filter(divisionName !='NY Security')
```






### Context 
I work as a Data Analyst at 32BJ Labor Union. Alot of our members are in NYC, 
but we do have members all down the East Coast (although I may analyze just the 
NYC ones for this one). Our Union Representatives hold Worksite Meetings at 
buildings to meet with Union Members, and we been tracking this for the past 10 
years. But recently in the past 2 years we've been tracking member attendance digitally
with our app (scanning member id cards), which we hope will be much more 
accurate. I will be just using this data from the past 2 years.
We are curious if having these meetings effects members union activity (event 
turnout etc), or political donations (ADF--American Dream Fund)

### Research question 

**You should phrase your research question in a way that matches up with the scope of inference your dataset allows for.**

Can we predict member turnout percentage for membership meetings?

turnout percentage ~ divisionName + membercount + ADFCount + city 


### Cases 

**What are the cases, and how many are there?**

There have been 577 worksite meetings in the past year in New York

### Data collection 

**Describe the method of data collection.**

Data is self-collected. I wrote the SQL query from our union database. The worksite meetings are inputted by our Field reps scanning member ids using our app.

### Type of study 

**What type of study is this (observational/experiment)?**

This is an observational study.

### Data Source 

**If you collected the data, state self-collected. If not, provide a citation/link.**

Self-collected.



### Describe your variables?

**Are they quantitative or qualitative**

quantitative

**If you are are running a regression or similar model, which one is your dependent variable?**

turnout percentage is my dependent variable


### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}
t <- sort(table(WorksiteMeetings$v3accountid), decreasing = TRUE)
```

**v3accountid: Number of times a site was visited through the year, most locations were only visited once.**
```{r}
describe(t)
```

**divisionName: Residential and Commercial buildings are our biggest divisions, this makes sense to me.**
```{r}
table(WorksiteMeetings$divisionName)
```
**v3fieldrepname**
```{r}
sort(table(WorksiteMeetings$v3fieldrepname),decreasing = TRUE)
```
**v3fieldrepname and divisionName**
```{r}
table(WorksiteMeetings$v3fieldrepname,WorksiteMeetings$divisionName)
```
**MemberCount is the roster size at that building. We represent lots of single doorman buildings, and smaller buildings with just a few cleaning staff, but we do have some larger 100+ person buildings**
```{r}
summary(WorksiteMeetings$memberCount)
```
```{r}
summary(WorksiteMeetings$ADFPercentage)
```

```{r}
ggplot(WorksiteMeetings, aes(x=(log(TurnoutPercentage)))) + geom_histogram()
```



```{r}
ggplot(WorksiteMeetings, aes(x=ADFcount,y=TurnoutPercentage)) + geom_point()
```

```{r}
ggplot(WorksiteMeetings, aes(x=ADFPercentage,y=TurnoutPercentage)) + geom_point()
```

```{r}
ggplot(WorksiteMeetings, aes(x=memberCount)) + geom_histogram()
```

**turnoutcount: Turnout at that individual event. Distribution is looking similar to building size, makes sense**
```{r}
describe(WorksiteMeetings$turnoutcount)
```
```{r}
ggplot(WorksiteMeetings, aes(x=turnoutcount)) + geom_histogram()
```

**TurnoutPercentage: percent of building that went to an event. There are alot of 1 person buildings, with 1 person attending for 100% attendance. I want to do something to account for turnout size, those bigger buildings with only 50% still have a huge number of ppl turn out. I should also probably use a log regression for my 0-1 scale.**
```{r}
ggplot(WorksiteMeetings, aes(x=TurnoutPercentage)) + geom_histogram(bins=30) +xlim(0,1.1)
```
**ADF Percentage, percentage of members paying into our political fund (we usually assume this people are more likely to be politically and union involved). happy to see a more normal distribution. again there are those 1 person buildings with that 1 member paying for 100% rate.**





```{r}
ggplot(WorksiteMeetings, aes(x=ADFPercentage)) + geom_histogram() +xlim(0,1.1)
```

```{r}
ggplot(WorksiteMeetings, aes(x=(turnoutcount))) + geom_histogram()
```

```{r}
summary(WorksiteMeetings$ADFcount)
```


```{r}
ggplot(WorksiteMeetings, aes(x=ADFcount)) + geom_histogram()
```
```{r}
ggplot(WorksiteMeetings, aes(x=ADFcount)) + geom_histogram()
```

**DateCreated: is the date the event happened. Unsure if i remember what exactly happening in April to cause the big spike in the spring, i know we had a few buildings try to forcefully switch to non-union staff this spring which caused some commotion, that could've been when reps were checking in more with those buildings?**
```{r}
ggplot(WorksiteMeetings, aes(x=datecreated)) + geom_histogram()
```


REGRESSION STUFF:

COUNT:
```{r}


turnout_count <- lm(turnoutcount ~ divisionName + memberCount + ADFcount, data = WorksiteMeetings)


summary(turnout_count)
```
```{r}
ggplot(data = turnout_count, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")
```
heavy tails, distribution is not normal
```{r}
ggplot(data = turnout_count, aes(sample = .resid)) +
  stat_qq() +
  stat_qq_line()
```
THIS ONE!!!!

```{r}
TESTING <- na.omit(WorksiteMeetings)
TESTING
```





```{r}


log_turnout_count2 <- glm(TurnoutPercentage ~ divisionName + memberCount + ADFcount, data = TESTING,family = "binomial"(link = "logit"), weights=turnoutcount)

summary(log_turnout_count2)


```

```{r}
1-pchisq(1131.38-952.04, 529-525)
```


Odds ratio (e^log-odds)
```{r}
exp(coefficients(log_turnout_count2))
```
membercount: negative sign indicates that the chance of turnout decreases as member count increases. 
ADFcount: negative sign indicates that the chance of turnout decreases as member count increases. 

()

I found the logistic regression interpretation a bit confusing, but these are my interpretation (while holding all other variables constant):

im not sure interpreting the intercept is that useful, assuming membercount is 0 and adf count is 0 

NY Residential buildings odds ratio of 0.49, suggests a 51% lower odds of turnout than a NY Commercial buildings

with the membercount odds ratio with 0.99, a increase of 1 more member at the buildings, would suggests a 1% lower odds of turnout 

similarly with ADFCount odds ratio being 0.99, a increase of 1 more person signing up for ADF at the buildings, would suggests a 1% lower odds of turnout 




```{r}

```


```{r}
plot_linear_assumption_check(log_turnout_count2)
```

```{r}
anova(log_turnout_count2)
```


```{r}
exp(-0.0071310)
```
```{r}
study$linear_resid <- resid(lm.out)
ggplot(study, aes(x = Hours, y = linear_resid)) +
    geom_point() +
    ggtitle('Residual Plot')
```


LINEAR MODEL:
```{r}
turnout <- lm(TurnoutPercentage ~ divisionName + memberCount + ADFPercentage, data = WorksiteMeetings)
summary(turnout)
```
```{r}
turnout <- lm(TurnoutPercentage ~ divisionName + memberCount, data = WorksiteMeetings)
summary(turnout)
```


""their variability changes as the proportion changes1 and as the sample size changes; consequently the results are heteroskedastic and generally nonlinear."" Note that count proportions are bounded between 0 and 1, and their variability changes as the proportion changes1 and as the sample size changes; consequently the results are heteroskedastic and generally nonlinear. (So linear regression might not be a great idea, for example.)

The usual approach for this sort of situation is some form of binomial regression, like logistic regression for example."

```{r}
log_turnout <- glm(TurnoutPercentage ~ divisionName + memberCount + ADFPercentage, data = WorksiteMeetings,family = "binomial")
summary(log_turnout)
```
google how to interpret log regressions

Member Count: for every 100 people at a location, we expect 1.1% less turnout?
ADFPercentage: for every 10% more of ADFPercentage you get at a building, we expect 15% more turnout at a worksite meeting


thats okay: "The intercept isn't significant because there isn't sufficient statistical evidence that it's different from zero"




```{r}
ggplot(data = turnout, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")
```
```{r}
ggplot(data = log_turnout, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals")
```


```{r}
ggplot(data = turnout, aes(x = .fitted, y = .resid)) +
  geom_jitter() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals") +
  geom_smooth() +
  geom_smooth(method="lm", se = FALSE, color="green")
```




```{r}
ggplot(data = turnout, aes(sample = .resid)) +
  stat_qq() +
  stat_qq_line()
```

```{r}
ggplot(data = log_turnout, aes(sample = .resid)) +
  stat_qq() +
  stat_qq_line()
```




```{r}

ggplot(WorksiteMeetings, aes(x=ADFcount, y=turnoutcount))  + geom_point()

```

```{r}
ggplot(WorksiteMeetings, aes(x=ADFPercentage, y=TurnoutPercentage))  + geom_point()

```



```{r}
ggplot(WorksiteMeetings, aes(x=turnoutcount, y=divisionName)) + geom_jitter()
```

```{r}
ggplot(WorksiteMeetings, aes(x=TurnoutPercentage, y=divisionName))  + geom_violin() + geom_jitter()

```


